\label[intro]\chap Introduction

The Real-Time Publish-Subscribe (\glref{RTPS})\cite[OMG:DDSI-RTPS22] is the protocol of Data Distribution Service (\glref{DDS})\cite[OMG:DDS] family, supporting Data-Centric Publish-Subscribe in real time and specifying communication in a decentralized network, where multiple nodes needs to send and/or receive data in real time. Specification of protocol is being developed by Object Management Group\cite[www:OMG] - international, open membership, not-for-profit technology standards consortium, since version 1.0 on February 2002 till version 2.2 on September 2014.

This thesis aims on upgrading \glref{ORTE} implementation of \glref{RTPS} protocol to be compatible with the latest standard version 2.2. The structure is as follows. In Chapter \ref[intro], there is an introduction to \glref{RTPS} and \glref{ORTE}. Chapter \ref[compare] compares implemented \glref{RTPS} 1.0 with the latest \glref{RTPS} 2.2, chapter \ref[upgrade] covers changes needed for compatibility with version 2.2 of the \glref{RTPS} protocol and chapter \ref[test] covers testing of new implementation of \glref{RTPS} protocol in \glref{ORTE}. In chapter \ref[shape], demo application of \glref{ORTE} called {\em Shape} for Android is introduced. The application was developed as part of familiarization with \glref{ORTE}, therefore \glref{ORTE} with \glref{RTPS} 1.0 implementation is used in this application. Security for \glref{DDS} is discussed in chapter \ref[sec].

\sec DDS

There are two main models used in Data Distribution Services. {\em Centralized} model, where single server for the whole network is needed and all communication goes throw it, introduces single point of failure. When the server is unreachable, the whole network is non-functional. By contrast, {\em decentralized} approach has no central server, no single point of failure. When one node of the network is non-functional, the rest of the network can continue in data transfers.

\sec DCPS

In the Data-Centric Publish-Subscribe network, data are sent by {\em Publishers} and received by {\em Subscribers}. Node can be {\em Publisher}, {\em Subscriber} or both and each node can be interested in different data, timing and reliability. Data-Centric Publish-Subscribe network is responsible for delivery of right data between right nodes with right parameters.

\rfc{obrazek}

\sec RTPS

Real-Time Publish-Subscribe is wire protocol developed to ensure interoperability between \glref{DDS} implementations. It has been designed to be fault tolerant (decentralized), scalable, tunable, with plug-and-play connectivity and ability of best-effort and reliable communication in real time applications.

\sec ORTE

Open Real-Time Ethernet (\glref{ORTE})\cite[FEE:ORTE] is the implementation of \glref{RTPS} 1.0. It's implemented in Application layer of UDP/IP stack, written in C, under open source license, with own API. Because there are no special requirements, it should be easy to port \glref{ORTE} to many platforms, where UDP/IP stack is implemented.

\sec Symbols

\makeglos

