\input ctustyle
\input opmac-bib
\input glos

\worktype [M/EN]
\faculty {F3}
\department {Department of measurement}

\title {Implementation of actual version of DDSI-RTPS protocol for distributed control in Ethernet network}
\titleCZ {Implementace aktuální verze protokolu DDSI-RTPS pro distribuované řízení v síti Ethernet}

\author {Jiri Hubacek}
\supervisor {Ing. Pavel Píša, Ph.D.}
%\authorinfo {jiri.hubacek@gmail.com}
\studyinfo {Open Informatics}

\date {January 2016}

\pagetwo {}

\abstractEN {
  This thesis describes process of upgrading Open Real-Time Ethernet implementation of Real-Time Publish-Subscribe protocol in order to be compatible with the current protocol version. Current version of the protocol is introduced, the original \glref{ORTE} implementation version of the protocol and the current version are compared and changes required for compatibility are discussed. Also demo application, future development and possibilities of security protocol extension are presented.
}
\abstractCZ {
  Tato práce popisuje proces aktualizace Open Real-Time Ethernet implementace Real-Time Publish-Subscribe protokolu za účelem zajištění kompatibility s aktuální verzí protokolu. Je představena aktuální verze protokolu, implementovaná verze protokolu a aktuální verze jsou porovnány a jsou diskutovány změny k zajištění kompatibility. Dále je prezentována demonstrační aplikace, budoucí vývoj a možnosti zajištění bezpečnosti.
}

\keywordsEN {
DDS, RTPS, ORTE, Ethernet, Real-Time
}
\keywordsCZ {
DDS, RTPS, ORTE, Ethernet, Real-Time
}

\thanks {
  Many thanks to Ing. Pavel Píša, Ph.D. for patience, comments and discussions and to Ing. Michal Sojka, Ph.D. for discussions.
}
\declaration {
  I, Jiri Hubacek, hereby declare that I am the author of the following text. All the citations and references are complete and properly named.
  \signature
}
\specification {\picw=\hsize \cinspic zadani.png}
%\draft
\makefront

% chapters here: \input some
\input intro
\input over
\input compare
\input upgrade
\input test
\input future
\input shape
\input sec
\input concl

\bibchap
\usebib/c (simple) mybase

\input app

\bye

